### 2019-03-12 paper.tex
Adding abstract, moving to ISWC template

\RequirePackage{amsmath}
\documentclass{llncs}
\usepackage[utf8]{inputenc}
\PassOptionsToPackage{table}{xcolor}
\pagestyle{headings}  % This option shows page numbers, it have to be removed in the final document.
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{scalerel}
\usepackage{comment}
\usepackage{framed}
\usepackage{listings}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\usepackage{balance}
\usepackage{stmaryrd}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{xspace}
\usepackage{url}
\usepackage{inconsolata}
\usepackage{etoolbox}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Quickly switch from conference to extended version
\providetoggle{conf}
\settoggle{conf}{false}

% text only to appear in extended version
\newcommand{\ever}[1]{\iftoggle{conf}{}{#1}}

% text only to appear in the conference version
\newcommand{\cver}[1]{\iftoggle{conf}{#1}{}}

% alternate text for both versions
\newcommand{\cever}[2]{\iftoggle{conf}{#1}{#2}}

% environment for long extended-only text
\iftoggle{conf}%
 {\excludecomment{exver}}%
 {\newenvironment{exver}}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{floatrow}
% Table float box with bottom caption, box width adjusted to content
\newfloatcommand{capbtabbox}{table}[][\FBwidth]

\newcommand{\ah}[1]{{\color{blue}\textsc{ah:} #1}}

\makeatletter
\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}%
	{1ex \@plus1ex \@minus.2ex}%
	{-1em}%
	{\normalfont\normalsize\itshape}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% SPARQL Listing style
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{listings}

\colorlet{punct}{red!60!black}
\definecolor{background}{HTML}{EEEEEE}
\definecolor{delim}{RGB}{120,20,40}
\definecolor{keyw}{RGB}{0,0,192}
\colorlet{numb}{magenta!60!black}

\lstdefinelanguage{sparql}{
	sensitive=false,
	extendedchars=true,
	literate={á}{{\'a}}1 {é}{{\'e}}1 {í}{{\'{\i}}}1 {ó}{{\'o}}1 {ú}{{\'u}}1
	{Á}{{\'A}}1 {É}{{\'E}}1 {Í}{{\'I}}1 {Ó}{{\'O}}1 {Ú}{{\'U}}1
	{ü}{{\"u}}1 {Ü}{{\"U}}1 {ñ}{{\~n}}1 {Ñ}{{\~N}}1 {¿}{{?``}}1 {¡}{{!``}}1
	{<}{{{\color{delim}<}}}{1}
	{>}{{{\color{delim}>}}}{1}
	{?}{{{\color{delim}?}}}{1}
	{*}{{{\color{delim}*}}}{1}
	{+}{{{\color{delim}+}}}{1}
	{/}{{{\color{delim}/}}}{1}
	{,}{{{\color{punct}{,}}}}{1}
	{;}{{{\color{punct}{;}}}}{1}
	{.}{{{\color{punct}{.}}}}{1}
	{:}{{{\color{punct}{:}}}}{1}
	{\{}{{{\color{delim}{\{}}}}{1} {\}}{{{\color{delim}{\}}}}}{1},
	morekeywords={ask,select,from,where,order,by,distinct,limit,offset,optional,union,filter,prefix,bound,desc,regex,str,group,not,exists,minus,service,certain,maybe}
}

\lstdefinestyle{sparqld}{
	basicstyle=\scriptsize\ttfamily,
	identifierstyle=\color{black},
	keywordstyle=\color{keyw}\bfseries,
	ndkeywordstyle=\color{greenCode}\bfseries,
	stringstyle=\color{ocherCode}\ttfamily,
	commentstyle=\color{darkgray}\ttfamily,
	language={sparql},
	tabsize=2,
	showtabs=false,
	showspaces=false,
	showstringspaces=false,
	extendedchars=true,
	escapechar=`,
	frame={single},
	breaklines=true,
	basewidth=0.5em,
	moredelim=[is][\color{magenta}]{~}{~},
	moredelim=**[is][\color{gray}]{£}{£},
	moredelim=**[is][\color{blue!50!black}]{$}{$},
	moredelim=[is][\color{orange!80!black}]{!}{!},
	moredelim=**[is][\color{green!50!black}]{¬}{¬},
	xleftmargin=2ex,
	xrightmargin=1ex,
	aboveskip=1.5ex,
	belowskip=1.5ex
}

\makeatletter
\newcommand{\sqbox}{%
	\collectbox{%
		\@tempdima=\dimexpr\width-\totalheight\relax
		\ifdim\@tempdima<\z@
		\fbox{\hbox{\hspace{-.5\@tempdima}\BOXCONTENT\hspace{-.5\@tempdima}}}%
		\else
		\ht\collectedbox=\dimexpr\ht\collectedbox+.5\@tempdima\relax
		\dp\collectedbox=\dimexpr\dp\collectedbox+.5\@tempdima\relax
		\fbox{\BOXCONTENT}%
		\fi
	}%
}
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% /SPARQL Listing style
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TiKz for RDF graphs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usetikzlibrary{shapes,arrows,positioning,fit,backgrounds,matrix,chains,scopes,calc}

\newcommand{\hsp}{\vphantom{Ag}}
\tikzset{
	std/.style={ 
		draw,
		circle,
		anchor=center,
		inner sep=0pt,
		minimum size=5pt
	},
	lab/.style={ 
		text centered,
		fill=white, 
		inner sep=0.7pt,
		font=\tt\small\hsp},
	iri/.style={
		draw=black!50!white, 
		rectangle,
		rounded corners,
		thick,
		text centered,
		top color=white, 
		bottom color=black!15,
		%opacity=0.7,
		%text opacity=1,
		font=\tt\small\hsp,
		anchor=center},
	lit/.style={
		draw=black!50!white, 
		rectangle,
		thick,
		text centered,
		top color=white, 
		bottom color=black!15, 
		anchor=center,
		%opacity=0.7,
		%text opacity=1,
		font=\tt\small\hsp},
	arrout/.style={
		->,
		-latex,
		%draw=black!50, 
		%fill=black!50,
		%thick,
		font=\tt\small\hsp},
	arrin/.style={
		<-,
		latex-,
		%draw=black!50, 
		%fill=black!50,
		%thick,
		font=\tt\small\hsp},
	arrinout/.style={
		<->,
		latex-latex,
		%draw=black!50, 
		%fill=black!50,
		%thick,
		font=\tt\small\hsp},
	arrstd/.style={
		-,
		%draw=black!50, 
		%fill=black!50,
		thick},
	dashmed/.style={
		-,
		%draw=black!50, 
		%fill=black!50,
		thick,
		dash pattern=on 3pt off 3pt,
		font=\tt\small\hsp},
	fade/.style={
		opacity=0.4,
		text opacity=0.4
	},
	fadet/.style={
		opacity=1,
		text opacity=0.4
	},
	every loop/.style={
		<-,
		latex-,
		fill=black!50,
		min distance=10mm,
		in=0,
		out=60,
		looseness=10,
		draw=black!50,
		thick,
		font=\tt\small\hsp
	},
	lean/.style={
		dotted,
		blue
	},
	leane/.style={
		dotted,
		blue
	},
	leanl/.style={
		text=blue
	}	
}

\newlength{\hgap}
\newlength{\vgap}
\setlength{\hgap}{2cm}
\setlength{\vgap}{1.2cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% /TiKz for RDF graphs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PGFplots
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{pgfplots}
\usepgfplotslibrary{groupplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=newest}
\usetikzlibrary{pgfplots.statistics}

\makeatletter
\pgfplotsset{
	boxplot prepared from table/.code={
		\def\tikz@plot@handler{\pgfplotsplothandlerboxplotprepared}%
		\pgfplotsset{
			/pgfplots/boxplot prepared from table/.cd,
			#1,
		}
	},
	/pgfplots/boxplot prepared from table/.cd,
	table/.code={\pgfplotstablecopy{#1}\to\boxplot@datatable},
	row/.initial=0,
	make style readable from table/.style={
		#1/.code={
			\pgfplotstablegetelem{\pgfkeysvalueof{/pgfplots/boxplot prepared from table/row}}{##1}\of\boxplot@datatable
			\pgfplotsset{boxplot/#1/.expand once={\pgfplotsretval}}
		}
	},
	make style readable from table=lower whisker,
	make style readable from table=upper whisker,
	make style readable from table=lower quartile,
	make style readable from table=upper quartile,
	make style readable from table=median,
}
\makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% /PGFplots
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{arydshln}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Hyperref
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{hyperref}
\definecolor{dark-blue}{rgb}{0.0,0.0,0.2}
\definecolor{dark-green}{rgb}{0.0,0.2,0.0}
\definecolor{dark-red}{rgb}{0.2,0.0,0.0}
\hypersetup{
	colorlinks, linkcolor={dark-red},
	citecolor={dark-green}, urlcolor={dark-blue},
	pdftitle={Canonicalisation of Monotone SPARQL Queries},    % title
	pdfauthor={Jaime Salas, Aidan Hogan},     % author
	pdfsubject={ISWC 2018},   % subject of the document
	pdfkeywords={sparql;} {canonicalisation;} {containment;} {query;} {equivalence;} {caching;}, % list of keywords
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% /Hyperref
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%% MACROS

\newcommand{\B}{\ensuremath{\mathbf{B}}\xspace}
\newcommand{\I}{\ensuremath{\mathbf{I}}\xspace}
\renewcommand{\L}{\ensuremath{\mathbf{L}}\xspace}
\newcommand{\V}{\ensuremath{\mathbf{V}}\xspace}

\newcommand{\cpx}[2]{\ensuremath{\textsc{#1-}\mathrm{#2}}\xspace}
\newcommand{\npc}{\cpx{NP}{complete}}
\newcommand{\nph}{\cpx{NP}{hard}}
\newcommand{\psc}{\cpx{PSpace}{complete}}
\newcommand{\psh}{\cpx{PSpace}{hard}}
\newcommand{\gic}{\cpx{GI}{complete}}
\newcommand{\gih}{\cpx{GI}{hard}}
\newcommand{\dpc}{\cpx{DP}{complete}}
\newcommand{\dph}{\cpx{DP}{hard}}
\newcommand{\conpc}{\cpx{coNP}{complete}}
\newcommand{\conph}{\cpx{coNP}{hard}}
\newcommand{\ptp}{\ensuremath{\Pi_2^P}}
\newcommand{\ptph}{\cpx{\ptp}{hard}}
\newcommand{\ptpc}{\cpx{\ptp}{complete}}

\newcommand{\tch}[1]{\textbf{#1}}
\newcommand{\rid}[1]{\textsc{#1}}
\newcommand{\ttl}[1]{\textsf{#1}}
\newcommand{\tid}[1]{\textsc{#1}}
%\newcommand{\dom}{\tid{dom}}
%\newcommand{\rng}{\tid{rng}}
%\newcommand{\sC}{\tid{sC}}
%\newcommand{\sP}{\tid{sP}}

\newcommand{\ssyn}[3]{[\ensuremath{#1\,\textsc{#2}\,#3}]}
\newcommand{\sand}[2]{\ssyn{#1}{and}{#2}}
\newcommand{\suni}[2]{\ssyn{#1}{union}{#2}}
\newcommand{\sopt}[2]{\ssyn{#1}{opt}{#2}}
\newcommand{\sfil}[2]{\ensuremath{\textsc{filter}_{#2}(#1)}}
\newcommand{\ssel}[2]{\ensuremath{\textsc{select}_{#2}(#1)}}
\newcommand{\sseld}[2]{\ensuremath{\textsc{select}^\Delta_{#2}(#1)}}
\newcommand{\dom}[1]{\ensuremath{\mathrm{dom}(#1)}}
\newcommand{\can}[1]{\ensuremath{\mathrm{can}(#1)}}
\newcommand{\com}[2]{\ensuremath{#1 \sim #2}}

%\newcommand{\hsc}[1]{{\footnotesize\MakeUppercase{#1}}}
\newcommand{\hsc}[1]{#1}

%\newcommand{\utp}[1]{\textsc{tp}{(#1)}}
\newcommand{\ufo}[1]{\textsf{\hsc{\upshape #1}}}
\newcommand{\uand}[1]{\ensuremath{\ufo{and}(#1)}}
\newcommand{\uuni}[1]{\ensuremath{\ufo{union}(#1)}}
\newcommand{\usel}[2]{\ensuremath{\ufo{select}_{#2}(#1)}}
\newcommand{\useld}[2]{\ensuremath{\ufo{select}^\Delta_{#2}(#1)}}

\newcommand{\uandn}{\ensuremath{\ufo{and}}}
\newcommand{\uunin}{\ensuremath{\ufo{union}}}

\newcommand{\bn}[1]{\texttt{\_:#1}}
\newcommand{\iri}[1]{\texttt{:#1}}
\newcommand{\var}[1]{\texttt{?#1}}

\newcommand{\ican}[1]{\ensuremath{\textsc{iCan}(#1)}}
\newcommand{\ecan}[1]{\ensuremath{\textsc{eCan}(#1)}}

\newcommand{\ev}[2]{\ensuremath{#1(#2)}}

\def\ojoin{\setbox0=\hbox{$\bowtie$}%
	\rule[0.18ex]{.25em}{.5pt}\llap{\rule[.9ex]{.25em}{.5pt}}}
\def\loj{\mathbin{\ojoin\mkern-5.8mu\bowtie}}

\newcommand{\da}{\ensuremath{:\nolinebreak\mkern-1.2mu\nolinebreak=}}

\newcommand{\qedr}{\begin{flushright}\qed\end{flushright}}

\newcommand{\yt}{\ding{51}}%
\newcommand{\nt}{\ding{55}}%

\newcommand{\para}[1]{\smallskip\noindent\textbf{#1:}}

\newcommand{\mq}{\textsc{mq}\xspace}
\newcommand{\mqs}{\textsc{mq}s\xspace}
\newcommand{\ucq}{\textsc{ucq}\xspace}
\newcommand{\ucqs}{\textsc{ucq}s\xspace}
\newcommand{\cq}{\textsc{cq}\xspace}
\newcommand{\cqs}{\textsc{cq}s\xspace}
\newcommand{\cuq}{\textsc{ucq}\xspace}
\newcommand{\cuqs}{\textsc{ucq}s\xspace}

%%%%%%%%

\graphicspath{ {images/} }

\begin{document}
\title{Versioned Queries in SPARQL}
\author{Ignacio Cuevas \and Aidan Hogan}

\institute{IMFD Chile \& Department of Computer Science, University of Chile}

\maketitle

\begin{abstract}
Although many prominent Linked Datasets are in a state of constant evolution, historical data are rarely published in a standard way: while dumps are sometimes provided of old versions of the dataset, there is no widely agreed-upon way for clients to quickly see historical changes for the description of a given resource,  or the results of a given SPARQL query, etc. In this paper, we focus on the problem of indexing historical versions of an RDF graph in a SPARQL engine and rewriting queries according to two versioning primitives: viewing results from a particular version, or viewing changes between two consecutive versions. Aiming to maintain compatibility with off-the-shelf SPARQL engines, we identify and compare several models by which historical versions of an RDF graph can be stored as named graphs, and discuss how base queries can be rewritten in each model per the two aforementioned primitives. We then present a performance comparison of these models with respect to X weekly versions of Wikidata and Y user-defined queries with versioning, using Virtuoso and Jena as reference SPARQL implementations. Our results show ...
\end{abstract}

\section{Introduction}
The Semantic Web~\footnote{\url{https://www.w3.org/2001/sw/wiki/Main_Page}} is an initiative of the World Wide Web Consortium, which envisions a ``Web of Data;'' that is, a web of \textit{Linked Data} where information has an standarized structure, making it easier for systems and machines to comunicate. In order to make such standarization possible, it becomes necessary to incorporate meta-data that describes the content, meaning and relationship of data in all databases.

The Resource Description Framework~\cite{key:rdfprimer11} (RDF for short) is a standard model for data interchange on the Web. Its design is based on ``triples,'' which represent connections on a graph, where every node is a ``resource'' with ``predicates'' as the edges connecting them. Both resources and predicates are formally defined, and a collection of rules allow for inferences about the relationships in the data.
Alongside RDF, the ontological languages RDFS~\cite{key:rdfschema11} and OWL~\cite{key:owl2primer} provide basic relation and type definitions, and SPARQL~\cite{key:sparql11} serves as RDF's query engine (similar to SQL on relational databases).

Currently, a large number of existing datasets are based on RDF due to its extense popularity, Wikidata being one the most notorious. Wikidata~\cite{VrandecicK14} is a massive dataset, with over 42 million items~\footnote{\url{https://www.wikidata.org/wiki/Wikidata:Statistics}}, which is based on all the available data of Wikipedia and structured on RDF.

Wikidata is in constant change, due to it being an open dataset that allows being edited by users. Furthermore, data is being added often, which makes Wikidata very dynamic in some areas. Given those circumstances, it can be said that Wikidata has many ``versions'' over time, where different data is available at different time intervals, making it possible to analyze its time component.

Dataset versioning in RDF/SPARQL is a research topic yet to be fully explored. While there are a number of implementations, most of them rely on specialized indices or SPARQL extensions. As such, it would be relevant to study the possibility of developing a versioning system and methodology for queries using only base SPARQL.

Such system could be used as version control for the dataset; administrators would be able to detect erroneous or malicious editions on data that doesn't normally change. On top of that, it would be possible to analyze the evolution of data, possibly predicting future changes. The system could also be used on datasets other than Wikidata, where data history may be poorly (or not at all) preserved.

\section{Related Work}
Gutierrez et al.~\cite{GutierrezHV07} describe a framework that allows time representation in RDF using annotations. An important distinction to make is that their work is focused on \textbf{labeling} over \textbf{versioning}; \textbf{labeling} means adding meta-data to query result validity, meanwhile \textbf{versioning} implies simultaneously keeping several versions of the same dataset. This work also formally defines labels and time intervals, but does not cover an implementation for either.

Zimmerman et. al~\cite{DBLP:journals/ws/ZimmermannLPS12} extend RDF to allow generalized annotations which can, in particular, be used to add validity time intervals to each triple.
\section{Preliminaries} % Maybe not
\section{(Method)} % Name of our method goes here
\section{Evaluation} % This two sections can
\section{Results}    % be merged for better flow
\section{Conclusion}
\newpage
\bibliographystyle{unsrt}
\bibliography{paper}
\end{document}