### 2020-09-09 query-rewrite.tex
added triple pattern rewriting for single version queries

Base pattern:
\begin{verbatim}
?s :p ?o
\end{verbatim}

\section{Single Version queries}
\subsection{Baseline}
The relevant version is select using a $\mathtt{FROM}$ or $\mathtt{GRAPH}$ statement.
\begin{verbatim}
FROM :i
\end{verbatim}
\vdots
\begin{verbatim}
?s :p ?o
\end{verbatim}
Alternatively:
\begin{verbatim}
GRAPH :i {?s :p ?o}
\end{verbatim}

\subsection{Intervals}
$\forall j \leq i, k \geq i$
\begin{verbatim}
FROM :j-k
\end{verbatim}
\vdots
\begin{verbatim}
?s :p ?o
\end{verbatim}
Alternatively $\mathtt{UNION}$ all :j-k where j $\leq$ i $\leq$ k in patterns such as:
\begin{verbatim}
GRAPH :j-k {?s :p ?o}
\end{verbatim}
\subsection{Absolute deltas}
The base version and ``positive'' deltas are combined as the default graph and the negative deltas are subtracted from the results.
\subsubsection{First version as base}
\begin{verbatim}
FROM :1
FROM :2-1
FROM NAMED :1-2
FROM :3-1
FROM NAMED :1-3
\end{verbatim}
\vdots
\begin{verbatim}
FROM :i-1
FROM NAMED 1-i
\end{verbatim}
\vdots
\begin{verbatim}
?s :p ?o
MINUS { GRAPH :1-2 {?s :p ?o} }
MINUS { GRAPH :1-3 {?s :p ?o} }
\end{verbatim}
\vdots
\begin{verbatim}
MINUS { GRAPH :1-i {?s :p ?o} }
\end{verbatim}

\subsubsection{Last version as base}
\begin{verbatim}
FROM :n
FROM :(n-1)-n
FROM NAMED :n-(n-1)
FROM :(n-2)-n
FROM NAMED :n-(n-2)
\end{verbatim}
\vdots
\begin{verbatim}
FROM :i-n
FROM NAMED n-i
\end{verbatim}
\vdots
\begin{verbatim}
?s :p ?o
MINUS { GRAPH :n-(n-1) {?s :p ?o} }
MINUS { GRAPH :n-(n-2) {?s :p ?o} }
\end{verbatim}
\vdots
\begin{verbatim}
MINUS { GRAPH :n-i {?s :p ?o} }
\end{verbatim}

\subsection{Sequential deltas}
The triple is searched in all deltas from the base up to the target. If the most recent delta is ``positive'', the match is preserved, otherwise it is discarded.
\subsubsection{First version as base}
\begin{verbatim}
{ SELECT  ?s ?o (MIN(?v0_) AS ?max_v0)
  WHERE
    { { { GRAPH ?g0
            { ?s  :p  ?o }
          ?g0  :direction  "forwards" .
          ?g0  :version  ?v0
        }
        BIND(str(?v0) AS ?v0_)
      }
      FILTER ( str(?v0) <= i )
    }
  GROUP BY ?s ?o
}
OPTIONAL
{ { SELECT  ?s ?o (MAX(?v1_) AS ?max_v1)
  WHERE
    { { { GRAPH ?g1
            { ?s  :p  ?o }
          ?g1  :direction  "backwards" .
          ?g1  :version  ?v1
        }
        BIND(str(?v1) AS ?v1_)
      }
      FILTER ( str(?v1) <= i )
    }
  GROUP BY ?s ?o
}
FILTER ( ( ! bound(?max_v1) ) || ( str(?max_v0) > str(?max_v1) ) )
}
\end{verbatim}
\subsubsection{Last version as base}
\begin{verbatim}
{ SELECT  ?s ?o (MIN(?v0_) AS ?max_v0)
  WHERE
    { { { GRAPH ?g0
            { ?s  <http://ex.org/p>  ?o }
          ?g0  :direction  "backwards" .
          ?g0  :version  ?v0
        }
        BIND(str(?v0) AS ?v0_)
      }
      FILTER ( str(?v0) >= i )
    }
  GROUP BY ?s ?o
}
OPTIONAL
{ { SELECT  ?s ?o (MIN(?v1_) AS ?max_v1)
  WHERE
    { { { GRAPH ?g1
            { ?s  :p  ?o }
          ?g1  :direction  "forwards" .
          ?g1  :version  ?v1
        }
        BIND(str(?v1) AS ?v1_)
      }
      FILTER ( str(?v1) >= i )
    }
  GROUP BY ?s ?o
}
FILTER ( ( ! bound(?max_v1) ) || ( str(?max_v0) < str(?max_v1) ) )
}
\end{verbatim}

\section{Version Delta queries}